---
layout: post
title: "Interactive Visualizations: Maps and Time Series"
date:   2015-1-26 20:49:52
authors: "Ben Best"
dateCreated:  2015-10-22
lastModified: 2015-10-23
category: visualization-workshop
tags: [module-5]
mainTag: GIS-Spatial-Data
description: "This post explains how to generate interactive map and time series visualizations."
image:
  feature: lidar_GrandMesa.png
  credit: LiDAR data collected over Grand Mesa, Colorado - National Ecological Observatory Network (NEON)
  creditlink: http://www.neoninc.org
permalink: /R/interactive-viz/
code1: /R/interactive.R
comments: true

---

<section id="table-of-contents" class="toc">
  <header>
    <h3>Contents</h3>
  </header>
<div id="drawer" markdown="1">
*  Auto generated table of contents
{:toc}
</div>
</section><!-- /#table-of-contents -->

##About
This activity will walk you through the fundamental principles of working 
with raster data in R.

**R Skill Level:** Intermediate - you've got the basics of `R` down.

<div id="objectives">

<h3>Goals / Objectives</h3>
After completing this activity, you will know:
<ol>
<li>What a raster dataset is and its fundamental attributes.</li>
<li>How to import rasters into `R` using the raster library.</li>
<li>How to perform raster calculations in `R`.</li>
</ol>

<h3>Things You'll Need To Complete This Lesson</h3>

<h3>R Libraries to Install:</h3>
<ul>
<li><strong>raster:</strong> <code> install.packages("raster")</code></li>
<li><strong>rgdal:</strong> <code> install.packages("rgdal")</code></li>

</ul>
<h4>Tools To Install</h4>

Please be sure you have the most current version of `R` and preferably
R studio to write your code.


<h4>Data to Download</h4>

Download the raster and *in situ* collected vegetation structure data:
<ul>
<li><a href="http://neonhighered.org/Data/LidarActivity/CHM_InSitu_Data.zip" class="btn btn-success"> DOWNLOAD Sample NEON LiDAR data in Raster Format & Vegetation Sampling Data</a></li>
<li><a href="{{ site.baseurl }}/data/rasterLayers_tif.zip" class="btn btn-success"> DOWNLOAD NEON imagery data (tiff format) California Domain 17 (D17)</a></li>
</ul>

<p>The LiDAR and imagery data used to create the rasters in this dataset were 
collected over the San Joaquin field site located in California (NEON Domain 17) 
and processed at <a href="http://www.neoninc.org" target="_blank" >NEON </a> 
headquarters. The entire dataset can be accessed by request from the NEON website.</p>  

<h4>Recommended Pre-Lesson Reading</h4>
<ul>
<li>
<a href="{{ site.baseurl }}/GIS-Spatial-Data/Working-With-Rasters/" target="_blank">
The Relationship Between Raster Resolution, Spatial extent & Number of Pixels - in R</a>
</li>
<li>
<a href="http://cran.r-project.org/web/packages/raster/raster.pdf" target="_blank">
Read more about the `raster` package in R.</a>
</li>
</ul>
</div>

#About Raster Data
Raster or "gridded" data are data that are saved in pixels. In the spatial world, 
each pixel represents an area on the Earth's surface. For example in the raster 
below, each pixel represents a particular land cover class that would be found in 
that location in the real world. 
<a href="{{ site.baseurl }}/GIS-Spatial-Data/Working-With-Rasters/"> More on 
rasters here</a>. 


# Group 5 - Visualization: Making Pretty Maps and Plots

**Notes** in Google Docs: [Module Five - Making Pretty Maps & Plots - Google Docs](https://docs.google.com/document/d/1g6E8bxHtT-GZFh3T99l4laZra7jHwM0y8h-3Slr-MAE/edit#)

**HTML** rendered of this Rmd: [bbest.github.io/NEON-DC-DataLesson-Hackathon/code/group05_visualization.html](http://bbest.github.io/NEON-DC-DataLesson-Hackathon/code/group05_visualization.html).


## Learning Objectives

After completing this activity, you will know:

- Visual Outputs, generic to time series or maps
    - Outputting to different formats: pdf, png w/ resolution. html/pdf.
    - post-process in Adobe Illustrator or Inkscape
    - Titles, axes labels. margins.
    - Legends
    - Tiling / Faceting.
    - Color ramps. Choosing color ramps for types of data: categorical, continuous. - divergent. colorblind. Best practice for this data. Link out to more.
- Production Quality Maps, ie Cartography
    - Scalebar, projection, N arrow. map-specific
    - Symbology. pts, lines.
- Interactive
    - web-based interactive plot using Javascript libraries 
    - specifics for R today, htmlwidgets R packages: leaflet, dygraphs
- Animation
    - existing raster time series animation of chm tower air shed
    - add time series next to it. point / pixel of individual stations + average. min/max/avg:

How to import rasters into `R` using the raster library.
How to perform raster calculations in `R`.



```{r setup}
suppressPackageStartupMessages({
  library(raster)       # work with rasters
  library(dplyr)        # work with data frames
  library(rgdal)        # read/write spatial files, gdal = geospatial data abstraction library
  library(xts)          # time series
  library(ggplot2)      # plotting
  library(readr)        # readr::read_csv() preferable to read.csv()
  library(knitr)        # knitting Rmarkdown to html
  library(animation)    # create animation ot the NDVI outputs
  library(scales)       # breaks and formatting for ggplot2
  library(lubridate)    # work with time
  library(leaflet)      # interactive maps htmlwidget
  library(RColorBrewer) # color ramps
  library(dygraphs)     # interactive time-series htmlwidget
  library(stringr)
})

#set the working directory
wd = '~/github/NEON-DC-DataLesson-Hackathon/code/1_WorkshopData'
setwd(wd)
opts_knit$set(root.dir=wd)
```

```{r}
dir = '~/github/NEON-DC-DataLesson-Hackathon/code/1_WorkshopData/Landsat_NDVI/HARV/2011/ndvi'

d = data_frame(
  fname = list.files(wd)) %>%
  mutate(
    sfx = str_replace(fname, '([0-9]+)_(.*)', '\\2'),
    jday = as.integer(str_extract(fname, '[0-9]+')))

sprintf('%03d_%s\n', d$jday, d$sfx)

# http://www.regexr.com/
```


```


## Quick Plot: Raster

```{r quick_raster}
# read raster
chm <- raster("NEON_RemoteSensing/HARV/CHM/HARV_chmCrop.tif")

# see metadata
chm

# quick plot
plot(chm, main="NEON Canopy Height Model (Tree Height)\nHarvard Forest")
```


## Production Quality Plot: Raster

```{r production_raster}
#customize legend, add units (m), remove x and y labels

```

## Interactive Plot: Raster

```{r interactive_raster}
# project if not already in mercator for leaflet
chm_mer = projectRasterForLeaflet(chm)

# get color palette
pal = colorNumeric(rev(brewer.pal(11, 'Spectral')), values(chm_mer), na.color = "transparent")

# output interactive plot
leaflet() %>%
  addTiles() %>%
  addProviderTiles("Stamen.TonerLite", options = providerTileOptions(noWrap = TRUE)) %>% 
  addRasterImage(chm_mer, colors=pal, project=F, opacity=0.8) %>%
  addLegend(pal=pal, position='bottomright', values=values(chm_mer), title='CHM')
```


## Production Plot: Time-Series

```{r production_timeseries}
harMet = read.csv('AtmosData/HARV/hf001-10-15min-m.csv')

#clean up dates
#remove the "T"
#harMet$datetime <- fixDate(harMet$datetime,"America/New_York")

# Replace T and Z with a space
harMet$datetime <- gsub("T|Z", " ", harMet$datetime)
  
#set the field to be a date field
harMet$datetime <- as.POSIXct(harMet$datetime,format = "%Y-%m-%d %H:%M", 
                          tz = "GMT")

#list of time zones
#https://en.wikipedia.org/wiki/List_of_tz_database_time_zones
#convert to local time for pretty plotting
attributes(harMet$datetime)$tzone <- "America/New_York"

#subset out some of the data - 2010-2013 
yr.09.11 <- subset(harMet, datetime >= as.POSIXct('2009-01-01 00:00') & datetime <=
as.POSIXct('2011-01-01 00:00'))

#as.Date("2006-02-01 00:00:00")
#plot Some Air Temperature Data
  
myPlot <- ggplot(yr.09.11,aes(datetime, airt)) +
           geom_point() +
           ggtitle("15 min Avg Air Temperature\nHarvard Forest") +
           theme(plot.title = element_text(lineheight=.8, face="bold",size = 20)) +
           theme(text = element_text(size=20)) +
           xlab("Time") + ylab("Mean Air Temperature")


#format x axis with dates
myPlot + scale_x_datetime(labels = date_format("%m/%d/%y"))
```


## Interactive Plot: Time-Series

```{r interactive_timeseries, eval=T}
# warning: resource intensive / time consuming to knit (~ 1 min)

# drop duplicate datetimes and assign rownames for later conversion to xts
ts = yr.09.11[!duplicated(as.character(yr.09.11$datetime)),]
row.names(ts) = as.character(ts$datetime)

date_window = row.names(ts)[c(
  which.max(row.names(ts) >= '2010-01-01 00:00:00'),
  which.min(row.names(ts) < '2011-01-01 00:00:00'))]

# plot
ts %>%
  select(airt) %>%
  as.xts() %>%
  dygraph() %>%
    dyRangeSelector(date_window)
```
